Codebook
========

Codebook generated on `r as.character(Sys.time())` 

Generated by `r R.version$version.string` on `r R.version$os`.

See `data/UCI HAR Dataset/features_info.txt` for original codebook

See `RunProject.html` for details on dataset creation.

Variable list and descriptions
------------------------------

Variable name | Description
--------------|------------
subject       | ID the subject who performed the activity for each window sample.
activityname  | Activity name
Mean-Time*    | Mean values from all samples for the subhect and activityname related to Time Domain Values
Mean-Freq*    | Mean values from all samples for the subhect and activityname related to Frequency Domain Values

Dataset structure
-----------------

```{r}
str(dtTidyAggregate)
```

Show a few rows of the dataset
------------------------------

```{r}
dtTidyAggregate
```

Summary of variables
--------------------

```{r}
summary(dtTidyAggregate)
```

Save to file
------------

Save data table objects to a tab-delimited text file called `AggregateDatasetHumanActivityRecognitionUsingSmartphones.txt`.

```{r save}
f <- file.path(path, "AggregateDatasetHumanActivityRecognitionUsingSmartphones.txt")
write.table(dtTidyAggregate, f, quote=FALSE, sep="\t", row.names=FALSE)
```